Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-12 volunteer-clouds-ieee-ci-si.tex
Few corrections

The sequence of events is as follows:
\begin{enumerate}
\item First the {\sf NodIO} process is started in Node.js. As part of the
initialization procedure the number of experiment is initialized and a log
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP
request and the page is loaded.
\item The main script renders the page and creates the worker instance. In
order to create a dedicated worker instance, another GET request is sent to the
server in order to receive the worker implementation script.
Each worker instance executes one or more evolutionary algorithms
{\em islands}. Each starts with random population and it is assigned a
universally unique  identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.
The initial parameters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following with the previously
specified parameters. Once the population has been initialized, the EA
generation loop is started. After {\em n} generations the state of the population
is evaluated. First the chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not a global best,
the EA loop must continue, but first a chromosome is randomly selected from
the shared population. Also the current generation number and the best
chromosome is sent to the client main thread in order to update the DOM to
display the current state of the island (described next). These messages are
all asynchronous, so even before a message is received by target the EA loop of
the worker could continue. If the best fitness is a global best the best
chromosome is sent to the server and the loop ends. A message is sent to the
main thread indicating that the best individual has been found.
\item The main thread receives messages from workers as events handled by
a callback function. When an iteration message is received, the page is
updated by displaying a dynamic plot of the generation number and best
fitness, and a representation of the chromosome is also generated.
\item When a global best is received from an island the current experiment
ends, the experiment number is incremented and the population array is reset.
\item When the main thread receives a message the a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the cost, only the parameters and population is reset and a new UUID
assigned.
\end{enumerate}
%
\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%

@@

However, it is interesting to note why this is so, and the first hint
is the inter-experiment correlation between the number of IPs in
successive experiments. While before it was an unremarkable 10\%, it is
now more than 43\%, making the number of IPs in contiguous experiments
highly correlated. The main reason for this is the setup in {\sf
  NodIO-W$^2$}, which restarts an island after a solution has been
found and also is running even if the tab is not on the
foreground. This means that volunteers can leave the experiment
running for as long as they want and they will be contributing to
experiment after experiment, unlike the previous version, where they
stopped contributing after the solution was found. This is also
reflected in the number of PUTs, which is the number of generations
divided by 100; more than 50\% of the volunteers run for more than
4000 generations. All things considered, this means than there are
many more computers running {\em simultaneously}, leading to this
almost 5-fold increase in running time.

@@

In order to establish a baseline performance, the evolutionary
algorithm was run in a desktop client-program written in JavaScript
using NodEO to solve the 40-trap function. The first experiments with
{\sf NodIO} proved that, although obtaining better performance than de
baseline was possible, it did not happen, mainly because of the
difficulty in carrying over volunteers to other experiments when the
one they were participating on was finished, shown in the low
correlation between the number of IPs in successive experiments. This
also resulted in a low number of generations allotted by users.

>>>>

The sequence of events is as follows:
\begin{enumerate}
\item First the {\sf NodIO} process is started in Node.js. As part of the
initialization procedure the number of experiment is initialized and a log
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP
request and the page is loaded.
\item The main script renders the page and creates the worker instance. In
order to create a dedicated worker instance, another GET request is sent to the
server in order to receive the worker implementation script.
Each worker instance executes over time one or more evolutionary algorithms
{\em islands}. Each starts with a random population and it is assigned a
universally unique  identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.
The initial parameters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following the previously
specified parameters. Once the population has been initialized, the EA
generation loop is started. After {\em n} generations the state of the population
is evaluated. First the chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not a global best,
the EA loop must continue, but first a chromosome is randomly selected from
the shared population. Also the current generation number and the best
chromosome is sent to the client main thread in order to update the DOM to
display the current state of the island (described next). These messages are
all asynchronous, so even before a message is received by the target, the EA loop of
the worker continues. If the best fitness is a global best the best
chromosome is sent to the server and the loop ends. A message is sent to the
main thread indicating that the best individual has been found.
\item The main thread receives messages from workers as events handled by
a callback function. When an iteration message is received, the page is
updated by displaying a dynamic plot of the generation number and best
fitness, and a representation of the chromosome is also generated.
\item When a global best is received from an island the current experiment
ends, the experiment number is incremented and the population array is reset.
\item When the main thread receives a message the a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the start-up cost, only the parameters and population are reset and a new UUID
assigned.
\end{enumerate}
%
\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%

@@

However, it is interesting to note why this is so, and the first hint
is the inter-experiment correlation between the number of IPs in
successive experiments. While before it was an unremarkable 10\%, it is
now more than 43\%, making the number of IPs in contiguous experiments
highly correlated. The main reason for this is the setup in {\sf
  NodIO-W$^2$}, which restarts an island after a solution has been
found and also is running even if the tab is not on the
foreground. This means that volunteers can leave the experiment
running for as long as they want and they will be contributing %borrado un to
experiment after experiment, unlike the previous version, where they
stopped contributing after the solution was found. This is also
reflected in the number of PUTs, which is the number of generations
divided by 100; more than 50\% of the volunteers run for more than
4000 generations. All things considered, this means than there are
many more computers running {\em simultaneously}, leading to this
almost 5-fold increase in running time.

@@

In order to establish a baseline performance, the evolutionary
algorithm was run in a desktop client-program written in JavaScript
using NodEO to solve the 40-trap function. The first experiments with
{\sf NodIO} proved that, although obtaining better performance than the
baseline was possible, it did not happen, mainly because of the
difficulty in carrying over volunteers to other experiments when the
one they were participating on was finished, shown in the low
correlation between the number of IPs in successive experiments. This
also resulted in a low number of generations allotted by users.
